---
title: "Reproducible Research Course Project 1"
author: "Shruti Sehgal"
date: "September 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the dataset
```{r, echo=TRUE, warning=FALSE}
library(ggplot2)
activity <- read.csv("E:/Coursera/Reproducible Research Course/repdata%2Fdata%2Factivity/activity.csv")
```

## Preprocessing
```{r, echo=TRUE, warning=FALSE}
activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

summary(activity)
```

## Mean total number of steps taken per day
```{r, echo=TRUE, warning=FALSE}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "red", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```


# Mean of the total number of steps taken per day:
```{r, echo=TRUE, warning=FALSE}
mean(activity_total_steps$steps)
```

# Median of the total number of steps taken per day:
```{r, echo=TRUE, warning=FALSE}
median(activity_total_steps$steps)
```

## The average daily activity pattern

# Time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r, echo=TRUE, warning=FALSE}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="red", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```

# 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps
```{r, echo=TRUE, warning=FALSE}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```


## Imputing missing values

# Total number of missing values in the dataset
```{r, echo=TRUE, warning=FALSE}
sum(is.na(activity$steps))
```

# Strategy for filling in all of the missing values in the dataset
```{r, echo=TRUE, warning=FALSE}
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]
```


# New dataset that is equal to the original dataset but with the missing data filled in
```{r, echo=TRUE, warning=FALSE}
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```

# Histogram of the total number of steps taken each day
```{r, echo=TRUE, warning=FALSE}
hist(total_steps_imputed$daily_steps, col = "red", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```

# The mean of the total number of steps taken per day
```{r, echo=TRUE, warning=FALSE}
mean(total_steps_imputed$daily_steps)

```
# The median of the total number of steps taken per day
```{r, echo=TRUE, warning=FALSE}
median(total_steps_imputed$daily_steps)
```
# After imputation, mean and median is the same.

# New factor variable in the dataset with two levels - "weekday" and "weekend"
```{r, echo=TRUE, warning=FALSE}
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })

```


#  Panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r, echo=TRUE, warning=FALSE}
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```


